---
title: "GLM for replicate"
output: html_document
date: "2023-03-23"
---

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse)
library(broom)
```

```{r}
for_colab_RA <- read_csv("https://raw.githubusercontent.com/tranJen/GLMNanoBiT_Honors_Thesis/main/GLM/Data_Akrong.csv")
for_colab_AM <- read_csv("https://raw.githubusercontent.com/tranJen/GLMNanoBiT_Honors_Thesis/main/GLM/Data_McGahan.csv")
```

#### *AM does not have D147E mutation but RA does.*

```{r}
df_AM <-filter(for_colab_AM[2:428,],library!=0)

df_RA  <-filter(for_colab_RA[2:429,],library!=0)
df_RA$B1 <- (df_RA$B1_1 + df_RA$B1_2 )/2
df_RA$B2 <- (df_RA$B2_1 + df_RA$B2_2 )/2
df_RA <- subset(df_RA, select = -c(B1_1,B1_2, B2_1, B2_2) )

df <- rbind(df_RA, df_AM)
df$pos <- as.character(df$pos)
df
```

# GLM for B1 > B2
```{r}
df$B1norm <- df$B1 / df$library
df$B2norm <- df$B2 / df$library
df$changeB2_B1 <- df$B2norm - df$B1norm
```
#### *p value is all high with 'binomial' but looks good with 'gaussian'*

## Gaussian
```{r}
gauss_model <- glm( changeB2_B1 ~ pos * mut , data = df, family = 'gaussian') 
#options(max.print=10000)    #set limit to print the output
summary(gauss_model)

#The log-likelihood is a measure of the log-probability of the observed data given the model parameters. Higher values of the log-likelihood indicate that the model is a better fit for the data.
logLik(gauss_model)
```
 
"Coefficients: (57 not defined because of singularities)" indicates that 57 of the covariates are a linear combination of some other of the covariates. In other words, it is because of singularity means that the variables are not linearly independent. If you remove the NA variables in the above summary, you will obtain the same result for the rest of the variables. This is because the information given by those variables is already contained in the other variables and thus redundant.

```{r}
fea <- names(coef(gauss_model))[44:length(names(coef(gauss_model)))]      #get features' names

#split the features' names (e.g from "pos137:mutA" to "137" and "A")
mut <- list()
pos <- list ()
for (i in fea){
              str_split <- strsplit(i, ":")
              pos <- append(pos, substr(str_split[[1]][1], 4,6))    # e.g extract 137
              mut <- append(mut, substr(str_split[[1]][2], 4,4))    # e.g extract A
}

est <- coef(gauss_model)[-1][43:length(coef(gauss_model)[-1])]          #get estimated contribution 
#p <- coef(summary(gauss_model))[-1,4]         #get p-value
p <- tidy(gauss_model)$p.value[44:483]        #to extract all including NA value
```

##### *white box is NA values. *

```{r}
#create data frame of pos, mut, estimated contribution, alpha
data_for_plot<- do.call(rbind, Map(data.frame, Pos=pos, Mut=mut, Est = est, p_value =p))
data_for_plot$alpha <- ifelse(data_for_plot$p_value<0.001,'***', ifelse(data_for_plot$p_value<0.01,'**', ifelse(data_for_plot$p_value<0.05,'*', ifelse(data_for_plot$p_value<0.1,'.', ' '))))
```

```{r}
#plot estimated contribution vs features
ggplot(data_for_plot, aes(x = Pos, y = Mut)) +
  geom_tile(aes(fill = Est)) + 
  geom_text(aes(label = alpha), size = 3)+
  scale_fill_gradient2(midpoint= 0, na.value = 'white',high = 'darkgoldenrod1 ', low = 'cyan4', mid ='aliceblue')+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 75, hjust=1), plot.subtitle = element_text(size = 10, color = "red", face = "bold"))+
  labs(y = "Mutation Amino Acid", x = "Enzyme Position", fill= 'Estimated
Contribution', subtitle = 'Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1')

```
### Remove unsignificant 
```{r}
data_for_plot <- data_for_plot[!(data_for_plot$Pos %in% c(137, 138, 139, 140, 141, 142,146,147,150,151,152,153,154,156,158)), ]
```

```{r}
heatmap <- ggplot(data_for_plot, aes(x = Pos, y = Mut)) +
  geom_tile(aes(fill = Est), width = 1, height = 1) +
  geom_text(aes(label = alpha), size = 3) +
  scale_fill_gradient2(midpoint= 0, na.value = 'white', high = 'darkgoldenrod1', low = 'cyan4', mid ='aliceblue') + 
  theme_classic() +
  theme(text =element_text(size = 10),
        # axis.text.x = element_text(angle = 75, hjust = 1), 
        plot.subtitle = element_text(size = 10, color = "red", face = "bold"), 
        aspect.ratio = 3/2,
        # legend.text = element_text(angle = 0, vjust = 1, hjust = 0.5),
        legend.title = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 10),
        legend.position =  c(1.2,0.5)) +
  guides(fill = guide_colorbar(title.position = "left", title.hjust = 0.5, label.position = "left"))+
  theme(legend.key.width = unit(0.25, "cm"))+
  labs(y = "Mutation Amino Acid", x = "Enzyme Position", fill= 'Estimated Contribution', subtitle = 'Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1')
heatmap
ggsave("heatmap.png", heatmap, width = 6, height = 5, units = "in")
```

